A community-maintained standard library of population genetic models

### 2019-12-20 popSimManu.tex
some reworking of intro

Assessing accuracy in population genetics is challenging in large part because
 ``ground-truth'' in population genetics comes from simulation,
not from empirical observations, and as a field there has been no agreement
on community standards or best practices. The general modus operandi to date has been to
introduce methods that are validated by a set of bespoke simulations produced by individual
groups. Beyond inefficiency from duplicated effort, such a situation leads
to decreased reproducibility and transparency across the entire field.
Here we describe \stdpopsim, a community-driven catalog
of standardized population genetics models spanning a variety of common study organisms.
This resource will empower the field to rigorously evaluate population genetic software,
%and select the best tool for a given task,
thereby substantially increasing the reliability of population genetic inferences.
%DRS: still not in love with this last sentence here, but tried to make it a little less sappy by adding some explanation.
%ADK: made some edits here
%RNG: Removed the sap. :-P Feel free to put it back if you like it better.
%JK: not crazy about it either. Making it shorter helps though?

It has been challenging to assess the degree to which modeling assumptions
and choices of data summaries can affect the ultimate results of population
genetic inference. Yet there
have been several clear examples of different methods yielding fundamentally
different conclusions. For example, Markovian coalescent methods applied to human genomes have
suggested large ancient ($>100,000$ years ago) ancestral population sizes and
bottlenecks that have not been detected by methods based on allele frequency spectra
\citep[see][]{beichman2017comparison}.
These distinct methods differ in how they model, summarize, and optimize fit to
genetic variation data, suggesting that such design choices can greatly affect the
performance of the inference. Furthermore, it is reasonable
to assume generally that specific methods will perform better than others under certain scenarios.
However, because these conditions are typically unknown, empirical researchers
lack principled guidelines for deciding which statistical method is best suited
to accurately answer their particular question. The need for empirical
guidance will only increase as researchers continue to generate more genomic
variation data from non-model taxa and wish to make inferences regarding
demography and selection.

One way to benchmark different methods for statistical inference in population
genetics is to apply the methods to simulated datasets where the true model
parameters are known. Methods that perform well will accurately infer these
parameters and will also provide reasonable measures of uncertainty of the
estimated parameters. Indeed, simulation studies are a standard ingredient in
most papers proposing new methods in population genetics. However, these
studies often use simulations of the particular models for which the method was designed,
which showcases the novel method rather than rigorously comparing performance to competing methods.
Of course, testing a new method in the situation for which it was designed is an important step,
but more rigorous testing is necessary for tools that are to be widely used in the field.
% Moreover researchers are generally forced to reimplement previously
% inferred scenarios independently, which makes good benchmarking
% a major additional time investment, and can lead to errors.
% Lastly, the degree to which different methods are compared can be quite variable.
Many more researchers in the broader field
use simulations to evaluate statistical power and perform other exploratory analyses.
The many steps of implementing a moderately complicated population genetics simulation,
including translating published demographic models into input for a simulator
and obtaining genetic maps,
present a major obstacle, especially to new researchers.

% One reason why simulation studies do not compare the performance of many methods
% across a wide range of situations
% is that simulation studies are challenging to carry out, for several reasons.
% First, while there are many published population genetic models available, their
% parameters are often not easily accessible, usually presented in supplemental
% tables. Translating these parameters into input for a given simulation engine can
% be a time-consuming and error prone process. In addition, papers often present many
% different models, and different authors may choose to select distinct models,
% making comparison difficult. While improvements in simulation software and ever
% faster computers have partially ameliorated this concern, generating the
% simulated datasets can still be a time and computing resource bottleneck.

For these reasons, we have generated a standardized, community-driven resource
for simulating published demographic models from a number of popular study systems.
This resource, which we call \stdpopsim, makes running
realistic simulations for population genetic analysis a simple matter of
choosing pre-implemented models from a community-maintained catalog.
The \stdpopsim catalog currently contains three organisms: humans,
\emph{Drosophila~melanogaster}, and \emph{Arabidopsis~thaliana}. For each
organism the catalog contains details on the physical organization (e.g., chromosome structure)
of their genomes, one or more genetic maps, default population-level parameters (mutation rate,
generation time) and one or more published demographic histories. Through
either a command line interface or a simple Python API, users can specify which
organism, genetic map, chromosome, and demographic history they are interested in simulating, and the
simulation output from their chosen model is returned.

The software has been developed by the PopSim Consortium using a
distributed open source model, with strong procedures in place
to continue its growth and maintain its quality.
Importantly, we have rigorous quality control methods to ensure implemented models are accurate
and have documented methods for others to contribute new modules.
We invite new collaborators to join our community.
Below we describe the resource and give
examples of how it can be used to benchmark demographic inference methods.

>>>>

Assessing the accuracy of inference methods for population genetics is
challenging in large part because the ``ground-truth'' in question
generally comes not from direct empirical observations, as the relevant
historical processes can rarely be observed, but instead from simulations.
Population genetic simulations are therefore critically important to the
field, yet there has been no systematic attempt to establish community
standards or best practices for executing them. Instead, the general modus
operandi to date has been for individual groups to validate their own
methods using bespoke simulations. 
Often these studies focus more on showcasing a novel method than on
rigorously comparing it with competing methods.
Moreover, this approach results in a great deal of duplicated effort across research
groups, and contributes to decreased reproducibility and transparency.
It is also a barrier to entry to the field, because new researchers can
struggle with 
the many steps involved
in implementing a state-of-the-art population genetics simulation,
including identifying appropriate demographic models from the literature, 
translating them into input for a simulator, obtaining appropriate genetic
maps, and choosing appropriate values for key population genetic parameters.

A related issue is that it has been challenging to assess the degree to which modeling assumptions
and choices of data summaries can affect population
genetic inferences. Yet there
are clear examples of different methods yielding fundamentally
different conclusions. For example, Markovian coalescent methods applied to human genomes have
suggested large ancient ($>100,000$ years ago) ancestral population sizes and
bottlenecks that have not been detected by methods based on allele frequency spectra
\citep[see][]{beichman2017comparison}.
These distinct methods differ in how they model, summarize, and optimize fit to
genetic variation data, suggesting that such design choices can greatly affect the
performance of the inference. Furthermore, some methods are likely to
perform better than others under certain scenarios, but 
researchers lack principled guidelines for selecting the best method for addressing 
their particular questions. The need for empirical
guidance will only increase as researchers 
seek to apply population genetic methods to a growing collection of non-model taxa.

For these reasons, we have generated a standardized, community-driven resource
for simulating published demographic models from a number of popular study systems.
This resource, which we call \stdpopsim, makes running
realistic simulations for population genetic analysis a simple matter of
choosing pre-implemented models from a community-maintained catalog, 
The \stdpopsim catalog currently contains three organisms: humans,
\emph{Drosophila~melanogaster}, and \emph{Arabidopsis~thaliana}. For each
organism the catalog contains details on the physical organization (e.g., chromosome structure)
of their genomes, one or more genetic maps, default population-level parameters (mutation rate,
generation time) and one or more published demographic histories. Through
either a command line interface or a simple Python API, users can specify which
organism, genetic map, chromosome, and demographic history they are interested in simulating, and the
simulation output from their chosen model is returned.
In this way, \stdpopsim will enable rigorous evaluation of population genetic software,
and contribute to increased reliability of population genetic inferences.

The \stdpopsim software has been developed by the PopSim Consortium using a
distributed open source model, with strong procedures in place
to continue its growth and maintain its quality.
Importantly, we have rigorous quality control methods to ensure implemented models are accurate
and have documented methods for others to contribute new modules.
We invite new collaborators to join our community.
Below we describe the resource and give
examples of how it can be used to benchmark demographic inference methods.
